Pyg4ometry: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-05-01 public/paper/pyg4ometry.tex
Update

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{xcolor}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

This paper describes such geometry creation and conversion package called {\em pyg4ometry}, implemented in {\em Python} and internally based on the Geant4 geometry 
description. The main features of  {\em pyg4ometry} are a python scripting API to rapidly design parametrised geometry, conversion to and from  FLUKA geometry 
descriptions, conversion from CAD formats  (STEP and IGES) based on FreeCAD \cite{FreeCAD} OpenCascade \cite{OpenCASCADE} and powerful geometry 
visualisation tools based on VTK \cite{VTK4}. The origin of pyg4ometry was a set of utilities to prepare geometry for an accelerator beam line simulation based on Geant4 
called BDSIM \colorbox{yellow}{[CITE]}. Accelerator physicists like specialists in other areas need a tool to quickly simulate beam losses, and the subsequent interaction of the charged particle 
beam with the accelerator material.

@@

There are some important issues which must be considered to perform an accurate conversion. 
A Geant4 logical volume solid might need to be created from a fluka region composing of many zones, 
so $R_1 = +z_1 | +z_2$, similarly a void is required to locate daughter volume placements which again 
might also be converted to a region of multiple zones $R_2= +z_3| z_4$. So creating the fluka region is then 
\begin{eqnarray}
R_1 - R_2 	& = & (+z_1 | +z_2) - ( +z_3 | +z_4) 			\\
			& = & (+z_1 - z_3 - z_4)  | (+z_2 - z_3 - z_4)	
\label{eqn:setDiff}
\end{eqnarray}
Geant4 has the boolean solids associated with difference, union and intersection, so in addition to Equation~\ref{eqn:setDiff}, both $R_1 \cup R_2$ and 
$R_1 \cap R_2$ is required in fluka notation, so 
\begin{eqnarray}
R_1 \cup R_2 	& = & (+z_1 | +z_2)  \cup ( +z_3 | +z_4) \\
			& = & +z_1 | +z_2 |  +z_3 | +z_4
\end{eqnarray}
and 
\begin{eqnarray}
R_1 \cap R_2 	& = & (+z_1 | +z_2) \cap ( +z_3 | +z_4) \\
			& = & +z_1 +z_3  | +z_1 +z_4 | +z_2 +z_3 | +z_2 +z_4
\end{eqnarray}

Fluka (apart from the {\it lattice} directive) has no sense of a volume hierarchy. Each body is placed with a 
translation, rotation, expansion geometry directives in global coordinates. A transformation from world coordinates
to a physical volume is built up by applying recursively daughter volume transformations and  this is used to place Fluka bodies. 
This  in practice is very similar to the procedure to create the VTK visualisation already described in Section~\ref{sec:visualisation}

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{siunitx}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{xcolor}
\usepackage{cleveref}

%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

This paper describes such geometry creation and conversion package called {\em pyg4ometry}, implemented in {\em Python} and internally based on the Geant4 geometry 
description. The main features of  {\em pyg4ometry} are a python scripting API to rapidly design parametrised geometry, conversion to and from  FLUKA geometry 
descriptions, conversion from CAD formats  (STEP and IGES) based on FreeCAD \cite{FreeCAD} OpenCascade \cite{OpenCASCADE} and powerful geometry 
visualisation tools based on VTK \cite{VTK4}. The origin of pyg4ometry was a set of utilities to prepare geometry for an accelerator beam line simulation based on Geant4 
called BDSIM \colorbox{yellow}{[CITE]}. Accelerator physicists like specialists in other areas need a tool to quickly simulate beam losses, and the subsequent interaction of 
the charged particle beam with the accelerator material. This paper describes version 1 of pyg4ometry, which is freely available as a git prepository and via pypi.

@@

There are some important issues which must be considered to perform an accurate conversion. 
A Geant4 logical volume solid might need to be created from a fluka region composing of many zones, 
so $R_1 = +z_1 | +z_2$, similarly a void is required to locate daughter volume placements which again 
might also be converted to a region of multiple zones $R_2= +z_3| z_4$. So creating the fluka region is then 
\begin{eqnarray}
R_1 - R_2 	& = & (+z_1 | +z_2) - ( +z_3 | +z_4) 			\\
			& = & (+z_1 - z_3 - z_4)  | (+z_2 - z_3 - z_4)	
\label{eqn:setDiff}
\end{eqnarray}
Geant4 has the boolean solids associated with difference, union and intersection, so in addition to Equation~\ref{eqn:setDiff}, both $R_1 \cup R_2$ and 
$R_1 \cap R_2$ is required in fluka notation, so 
\begin{eqnarray}
R_1 \cup R_2 	& = & (+z_1 | +z_2)  \cup ( +z_3 | +z_4) \\
			& = & +z_1 | +z_2 |  +z_3 | +z_4
\label{eqn:setUnion}
\end{eqnarray}
and 
\begin{eqnarray}
R_1 \cap R_2 	& = & (+z_1 | +z_2) \cap ( +z_3 | +z_4) \\
			& = & +z_1 +z_3  | +z_1 +z_4 | +z_2 +z_3 | +z_2 +z_4
\label{eqn:setIntersection}
\end{eqnarray}

Fluka (apart from the {\it lattice} directive) has no sense of a volume hierarchy. Each body is placed with a 
translation, rotation, expansion geometry directives in global coordinates. A transformation from world coordinates
to a physical volume is built up by applying recursively daughter volume transformations and  this is used to place Fluka bodies. 
This  in practice is very similar to the procedure to create the VTK visualisation already described in Section~\ref{sec:visualisation}.

In Fluka every single space point needs to be associated with one and only one region. There is problem when converting
geant4 logical Volumes to fluka regions, as the logical volume outer solid $S_{\rm logical}$ need to have the daughter 
solids $S_{{\rm daugher},i}$ subtracted. So a solid which can be converted to a region $S_{\rm region}$ is then 
\begin{equation}
S_{\rm region} =  S_{\rm logical} - S_{\rm daughter,1} - S_{\rm daughter,2} - S_{\rm daughter,3} \ldots 
\label{eqn:logical}
\end{equation} 
If a logical volume has  number of daughter volumes which are also possibly boolean solids, then computing 
$S_{\rm region}$ can become very complex because of Equations \ref{eqn:setDiff}, \ref{eqn:setUnion} and 
\ref{eqn:setIntersection}.
